### 2015-07-27 text/ms.Rmd
Trivial tweaks
Incorporate a round of co-author comments

<!--TODO: Jessicas Comments: 
15. Line 255 from pdf - CMSY predictions were approximately linear  (Fig S4) - Do you mean Fig S4a? Maybe I don't know how to interpret these graphs - I don't see how you can extract that info from the CMSY graph. I don't really know wht you mean by 'high and low prediction clusters'.
16. line 258-59 - if mPRM and SSCOM are high, then true status tends to be high. Is there another example that shows how covariates are exploited. The example shown is convincing in Fig S5, but not in Fig S6l, because the predicted B/Bmsy is only just above (light pink shading). though ths may not matter. you can ignore this comment if you want.
18. line 314-15 need a reference showing that SSCOM contributed little to status from GBM. e.g. which figure shows this?
19. line 334 - should this be "effort dynamics at low biomasses in COM-SIR", not SSCOM? 
20. I said that i'd help with the last paragraph about other uses for ensemble methos. If there is space within the word limit, could we add just one other potential use? Are the references listed in 350-51 using ensemble models or do they simply provide examples of where ensembles could be used. Just check my addition - and delete it if you aren't happy with it. I don't know how to insert references - i'll leave that up to you. 
-->

@@

\noindent
Sean C. Anderson^1\*^, 
Andrew B. Cooper^1^,
Olaf P. Jensen^4^, 
C\'{o}il\'{i}n Minto^6^, 
James T. Thorson^11^,
Jessica C. Walsh^1^,
Jamie Afflerbach^2^, 
Mark Dickey-Collas^3,12^, 
Kristin M. Kleisner^5^, 
Catherine Longo^2^, 
Giacomo Chato Osio^7^, 
Dan Ovando^8^, 
Andrew A. Rosenberg^9^, 
Elizabeth R. Selig^10^

@@

\noindent
^5^NOAA/Northeast Fisheries Science Center, Ecosystem Assessment Program, 
166 Water St., Woods Hole, MA, 02543, USA

@@

Ecological resource managers are often faced with reconciling multiple potentially conflicting estimates of population status and trajectory. One solution is to average these predictions in an ensemble, but this approach ignores covariance between models and may not optimally reduce bias and improve accuracy. Superensemble models, commonly used for climate and weather forecasting, may provide a superior solution. Superensembles use the predictions from multiple models as covariates in a new statistical model. Here we evaluate the potential for ensemble and superensemble models ('ensemble methods') to improve estimates of fish population status and trajectory. We fit four data-limited models of population and exploitation dynamics to marine fish stocks and combine their estimates of the mean and slope of biomass at maximum sustainable yield ($B/B_\mathrm{MSY}$) with an ensemble average and three superensembles: a linear model, a random forest, and a boosted regression tree. We selected data-limited models for their applicability across the majority of fish stocks world wide. We built our models on a simulated dataset of 5760 stocks and tested them with cross-validation and a global database of `r ram_stocks_n` stock assessments that have also been assessed with more data-intensive stock assessment models. We find that ensemble methods substantially improve estimates of population status and trajectory. Machine-learning superensembles performed the best for population status: accuracy improved `r mean_sim$mach_vs_ind_mare_fold`%, rank-order correlation between predicted and true status improved from `r mean_sim$ind_corr_range` to `r mean_sim$mach_corr_range`, and bias (median proportional error) declined from `r mean_sim$ind_mre_range` to `r mean_sim$mach_mre_range`. We found similar improvements when predicting the trajectory of status and when applying the simulation-trained superensembles to catch data for global fish stocks. Ensemble methods have the potential to improve estimates of population status and trajectory across many taxa beyond fish. However, they must be carefully tested, formed from a diverse set of accurate models, and built on a dataset sufficiently representative of the populations to which they are applied.

@@

Population status and trajectory are two of the most fundamental values to quantify in conservation biology [e.g. @mace2008; @iucn2015]. These properties are key to everything from global assessments of extinction risk [e.g. @stuart2004; @schipper2008] to assessing the sustainability of hunting and fishing exploitation levels [e.g. @worm2009; @juan-jorda2011]. However, conservation and ecological resource managers are often faced with reconciling multiple uncertain and potentially conflicting estimates of status and trajectory [e.g. @brodziak2010; @branch2011; @deroba2015]. Some models may suggest a population is at risk and declining in abundance while others may suggest it is not at risk and stable.

@@

In fisheries science, the commonly used operational model for determining status and trajectory of exploited fish populations are stock assessments that incorporate all appropriate data (e.g. catches, size and age distributions, surveys, and tagging information) to quantify values such as the biomass of a stock that can produce maximum sustainable yield ($B/B_\mathrm{MSY}$). However, the broad range of data required to conduct these stock assessments are not available for the majority of fish populations, including those of conservation concern and of economic interest to fisheries [@fao2014]. Therefore, a number of models have been proposed to assess $B/B_\mathrm{MSY}$ based on the limited data available for the majority of fish stocks: catches and a basic understanding of population productivity [e.g. @vasconcellos2005; @martell2013]. Recently, @rosenberg2014 investigated the performance of four data-limited models through a large-scale simulation experiment. The four models frequently disagreed about population status and any one model had relatively poor predictive ability on average (e.g. Fig. \ref{motivate}).

Here, we estimate population status and trajectory of exploited fish populations using ensembles and superensembles (collectively 'ensemble methods') of these four data-limited models. We apply a variety of approaches to both simulated and real-world fish stocks and compare their predictive performance. We find that ensembles, and in particular superensembles, generally improved predictive performance of status and trajectory of fish populations over any single model---especially when considered across multiple dimensions of predictive ability. Furthermore, the output from superensembles can aid mechanistic understanding of how individual models of status and trajectory perform.

@@

We first developed and tested ensemble methods with a fully factorial simulated dataset [@rosenberg2014] with precisely known status. These published simulations included the following scenarios: three fish life histories: small pelagic, demersal, and large pelagic; three levels of how depleted biomass was at the start of the dataset compared to carrying capacity: 100%, 70%, and 40% of carrying capacity; and four harvest dynamics: a constant harvest rate, a harvest rate coupled with biomass to mimic an open-access single-species fishery, a scenario where harvest rate increases continuously, and a 'roller-coaster' scenario where the harvest rate increases and then decreases. Process noise (i.e. variability in population dynamics) was introduced to the models through two levels of multiplicative recruitment variability: $N(0, 0.2)$ and $N(0, 0.6)$, and was either uncorrelated or had a first-order autoregressive correlation of $0.6$. The simulation also included a scenario with multiplicative observation error around catch at $N(0, 0.2)$ and one without observation error. 
<!--TODO: make last sentence clearer
4. First paragraph in Methods describing simulations line 113 in pdf - you may want to describe the recrutiment variability a bit more. I know its not an important detail, but because its in the first paragraph of methods, it might turn people off reading further. Most people wouldn't understand what 'multiplcative' (including me) is or may not know what 'process error/noise' refers to. 
5. First paragraph in Methods last sentence - needed a bit of rewording because as was, i got quite confused. hopefully this new sentence makes sense and doesnt change the meaning. 
-->

@@

We also tested our ensemble methods on the RAM Legacy Stock Assessment Database [@ricard2012]. Our analysis of the stock-assessment database was based on version XX downloaded on XX (TODO). After removing stocks for which the models described below did not converge (TODO: add number), this database included `r ram_stocks_n` stocks.

@@

We fit four individual data-limited models to estimate $B/B_\mathrm{MSY}$. Three of the models are mechanistic and based generally on Schaefer biomass dynamics [@schaefer1954] of the form

@@

\noindent
where $\hat{B}_{t+1}$ represents predicted biomass at time $t$ plus one year, $B_t$ represents biomass at time $t$, $r$ represents the intrinsic population growth rate, $B_0$ represents unfished biomass or carrying capacity $K$, and $C$ represents catch. @rosenberg2014 provide a full summary of these four methods and code to fit all the models is available in an accompanying package `datalimited` (TODO: make public?) for the statistical software \textsf{R}. In summary:

* *CMSY* (catch-MSY) implements a stock-reduction analysis with Schaefer biomass dynamics [@martell2013]. It requires a prior distribution for $r$ and assigns a prior to the relative proportion of biomass at the end of the time series compared to unfished biomass (depletion) based on the percentage of maximum catch at the end of the time series. Our version of the model was modified from @martell2013 to generate biomass trends from all viable $r$-$K$ pairs and produce an estimate of $B/B_\mathrm{MSY}$ from the median trend.

* *COM-SIR* (catch-only-model with sample-importance-resampling) is a coupled harvest-dynamics model [@vasconcellos2005]. Biomass is assumed to follow a Schaefer model and harvest dynamics are assumed to follow a logistic model. The model is fit with a sample-importance-sampling algorithm [@rosenberg2014].

@@

* *mPRM* (modified panel regression model) is a modified version of the panel-regression model from @costello2012. Unlike the other models, mPRM is empirical and not mechanistic --- it uses the RAM Legacy Stock Assessment database to fit a regression model to a series of characteristics of the catch time series and stock with stock-assessed $B/B_\mathrm{MSY}$ as the response. The model used in this paper is modified from the original in that it condenses the life-history categories into three categories to match the simulated dataset, removes the maximum catch predictor since the absolute catch in the simulated dataset is arbitrary, and does not implement the bias correction needed in @costello2012 for deriving aggregate estimates of median status across multiple stocks.

@@

<!--TODO Kristin and Mark: It might be a good idea to add a figure that illustrates the spectrum of ensemble/superensembles that are applied here and what is included (e.g., non-linear relationships, etc.). This will help the reader understand the progression in complexity of the ensemble approaches I think.-->
Superensemble models allow us to incorporate additional covariates and potentially leverage interactions between these covariates and individual model predictions. Additional covariates could be, for example, life-history characteristics, information on exploitation patterns, or statistical properties of the data. For simplicity, and to allow us to apply models developed with the simulated dataset to the real-world dataset, we added only one set of additional covariates: spectral properties of the catch time series. Spectral analysis decomposes a time series into the frequency domain and provides a means of describing the statistical properties of the catch series that is independent of time series length and absolute magnitude of catch. We fit spectral models to the scaled catch time series (catch divided by maximum catch) with the `spec.ar` function in \textsf{R} and recorded representative short- and long-term spectral densities at frequencies of 0.20 and 0.05, which correspond to 5- and 20-year cycles.

@@

The individual models we seek to combine with superensembles provide time series of $B/B_\mathrm{MSY}$. Therefore, we can use superensembles to estimate multiple properties of these status time series. Here, we focus on two properties: the mean and slope of $B/B_\mathrm{MSY}$ in the last 5 years. Together, these quantities address the recent state and trajectory of status, which are both of management and conservation interest [e.g. @mace2008]. To avoid undue influence of the end points of the time series on the calculated slope, we measured the slope as the Theil-Sen estimator of median slope [@theil1950].

@@

A critical component to any predictive modelling exercise is to evaluate the performance of a model on new data [@hastie2009]. We used repeated three-fold cross validation to test predictive performance: we randomly divided the dataset into three sets, built superensemble models on two-thirds of the data, and evaluated predictive performance on the remaining third. We repeated this across each of the three splits and then repeated the whole procedure 50 times to account for bias that may result from any one set of validation splits. Since the dynamics of simulated populations differing only in the stochastic draw of random values are likely similar, we grouped these stocks (10 per factorial cell) in the cross-validation process into either the training or testing split. When applied to the RAM Legacy Stock Assessment database, we used superensembles trained on the entire simulation dataset. However, since mPRM is built on the same stock assessment database, we refit mPRM on each training split.

Predictive performance can be evaluated with metrics that represent a variety of modelling goals. For continuous response variables such as the mean and slope of population status, performance metrics often measure some form of bias, precision, accuracy (a combination of bias and precision), or the ability to correctly rank or correlate across populations [e.g. @walther2005]. Here, we measure proportional error, defined as $(\hat{b} - b)/b$, where $\hat{b}$ and $b$ represent estimated and 'true' (or stock-assessed) mean or slope of $B/B_\mathrm{MSY}$. We calculated median proportional error to measure bias, median absolute proportional error to measure accuracy, and Spearman's rank-order correlation between predicted and 'true' values to measure the ability to correctly rank populations. We measured the ability for each model to correctly place stocks into categories of above or below $B/B_\mathrm{MSY} = 1$ at any decision threshold by plotting receiver operating characteristic curves. Code to reproduce our analysis is available at <https://github.com/datalimited/ensembles> (TODO: make public).

@@

Applied to the simulated dataset of known status, the individual models had variable success at recovering the mean (status) and slope (trajectory) of $B/B_\mathrm{MSY}$ in the last five years. All models exhibited a high degree of scatter around the one-to-one line of perfect status prediction (Fig. \ref{hexagon}). CMSY exhibited bimodal predictions of status (Fig. \ref{hexagon}a), but had the best rank-order correlation and accuracy scores (Fig. \ref{performance}a). COM-SIR and SSCOM both correctly identified a number of low status stocks, but frequently predicted a high status when status was in fact low (Fig. \ref{hexagon}b, c). mPRM had relatively poor ability to predict status for the simulated dataset (Fig. \ref{hexagon}d). There was generally little correlation between true and predicted trajectory for any of the individual models (rank-order correlation = `r range_slope_corr_non_sscom`) besides SSCOM (correlation = `r slope_corr_sscom`; Figs \ref{scatter-sim-slope}a--d).

@@

@@

@@

@@

Certain conditions will make some ensemble models more effective than others. First, ensembles will be most effective when they are comprised of diverse individual models that choose different structural model forms, explore contrasting but plausible ranges of parameter values, or make errors in uncorrelated ways [@ali1996; @dietterich2000; @tebaldi2007]. Such individual models would be expected to perform well in different conditions and an ensemble model can exploit the best predictive performance of each. Second, ensemble models will be most effective when they are not overfit to the training dataset. Cross-validation testing [@caruana2004; @hastie2009] and methods that are robust to overfitting such as random forests [@breiman2001], may help avoid overfitting ensemble models. We note that our simplest ensemble model, an average of individual model predictions, performed approximately as well as complex machine learning models when we trained our superensembles on the simulation dataset and tested them on a separate 'real' dataset (i.e. the RAM Legacy Stock Assessment database, Fig. \ref{performance}b). Third, ensemble models will be most effective when they are trained on data that are representative of the dataset of interest [@knutti2009; @weigel2010]. Cross-validation within a training dataset will provide an optimistically biased impression of predictive performance if the training dataset fundamentally differs from the dataset of interest [@hastie2009].

Multi-model inference in the form of coefficient averaging weighted by information theoretics such as the Akaike Information Criterion (AIC) is a common analytical approach in ecology [@burnham2002; @johnson2004; @grueber2011]. The ensemble methods described in this paper share similarities with coefficient averaging but differ in other important ways. Ensemble methods and coefficient averaging share the long-held notion that multiple working hypotheses can contribute useful information for inference [@chamberlin1890]. A fundamental difference is that coefficient averaging focuses on averaging *coefficients* whereas ensembles instead average *predictions*. While the outcome will be the same in some cases, ensembles provide a more general purpose tool: they do not require information theoretics and they can combine different types of models (e.g. parametric and non-parametric models or frequentist and Bayesian predictions). Furthermore, superensembles extend these benefits by allowing model predictions to be combined via non-linear functions that are tuned to known data.

A strength of superensembles is that they can be tailored to predict specific response variables. For example, we built separate superensemble models of mean $B/B_\mathrm{MSY}$ and the slope of $B/B_\mathrm{MSY}$. The same set of model weights or non-linear relationships need not hold across different response variables. For instance, SSCOM contributed little to the GBM superensemble estimate of status at higher levels of predicted $B/B_\mathrm{MSY}$, but contributed strongly to estimates of trajectory (Fig. \ref{partial-sim-slope}). Formally, fitting superensemble models to specific quantities of interest provides an additional calibration step [@rykiel1996]. Individual models are typically calibrated (i.e. parameters are estimated) based on a single response variable; however, this model might then provide a biased or inaccurate estimate of other quantities of interest. Ensemble methods provide a final calibration step to a quantity of interest. This ensemble calibration could further include a loss function tailored to the goals of the model, say placing greater weight on accuracy at low status levels than higher values. Conversely, because superensembles are tailored to a specific response and loss function, superensembles force a modeller to choose an operational purpose for their model up front [*sensu* @dickey-collas2014].

>>>>

<!--TODO: Jessicas Comments: 
TODO - should the table go in the main text, new fig, grid image?
15. Line 255 from pdf - CMSY predictions were approximately linear  (Fig S4) - Do you mean Fig S4a? Maybe I don't know how to interpret these graphs - I don't see how you can extract that info from the CMSY graph. I don't really know wht you mean by 'high and low prediction clusters'.
16. line 258-59 - if mPRM and SSCOM are high, then true status tends to be high. Is there another example that shows how covariates are exploited. The example shown is convincing in Fig S5, but not in Fig S6l, because the predicted B/Bmsy is only just above (light pink shading). though ths may not matter. you can ignore this comment if you want.
18. line 314-15 need a reference showing that SSCOM contributed little to status from GBM. e.g. which figure shows this?
19. line 334 - should this be "effort dynamics at low biomasses in COM-SIR", not SSCOM? 
20. I said that i'd help with the last paragraph about other uses for ensemble methos. If there is space within the word limit, could we add just one other potential use? Are the references listed in 350-51 using ensemble models or do they simply provide examples of where ensembles could be used. Just check my addition - and delete it if you aren't happy with it. I don't know how to insert references - i'll leave that up to you. 
-->

@@

\noindent
Sean C. Anderson^1\*^, 
Andrew B. Cooper^1^,
Olaf P. Jensen^4^, 
C\'{o}il\'{i}n Minto^6^, 
James T. Thorson^11^,
Jessica C. Walsh^1^,
Jamie Afflerbach^2^, 
Mark Dickey-Collas^3,12^, 
Kristin M. Kleisner^5^, 
Catherine Longo^2^, 
Giacomo Chato Osio^7^, 
Daniel Ovando^8^, 
Andrew A. Rosenberg^9^, 
Elizabeth R. Selig^10^

@@

\noindent
^5^National Marine Fisheries Service, National Oceanographic and Atmospheric
Administration, Northeast Fisheries Science Center, Ecosystem Assessment
Program, 166 Water St., Woods Hole, MA, 02543, USA

@@

Ecological resource managers are often faced with reconciling multiple potentially conflicting estimates of population status and trajectory. One solution is to average these predictions in an ensemble, but this approach ignores covariance between models and may not optimally reduce bias and improve accuracy. Superensemble models, commonly used for climate and weather forecasting, may provide a superior solution. Superensembles use the predictions from multiple models as covariates in a new statistical model. Here we evaluate the potential for ensemble and superensemble models ('ensemble methods') to improve estimates of fish population status and trajectory. We fit four data-limited models of population and exploitation dynamics to marine fish stocks and combine their estimates of the mean and recent trends in biomass relative to the equilibrium biomass at maximum sustainable yield ($B/B_\mathrm{MSY}$) with an ensemble average and three superensembles: a linear model, a random forest, and a boosted regression tree. We selected data-limited models for minimal data requirements and thus wide applicability across the majority of the world's fish stocks. We built our models on a simulated dataset of 5760 stocks and tested them with cross-validation and against a global database of `r ram_stocks_n` stock assessments that incorporated data-intensive stock assessment models. We found that ensemble methods substantially improved estimates of population status and trajectory. Machine-learning superensembles performed the best for estimating population status: accuracy improved `r mean_sim$mach_vs_ind_mare_fold`%, rank-order correlation between predicted and true status improved from `r mean_sim$ind_corr_range` to `r mean_sim$mach_corr_range`, and bias (median proportional error) declined from `r mean_sim$ind_mre_range` to `r mean_sim$mach_mre_range`. We found similar improvements when predicting the trajectory of status and when applying the simulation-trained superensembles to catch data for global fish stocks. Ensemble methods have the potential to improve estimates of population status and trends for a wide variety of plant and animal species. However, they must be carefully tested, formed from a diverse set of accurate models, and built on a dataset sufficiently representative of the populations to which they are applied.

@@

Population status and trajectory are two of the most fundamental values to quantify in conservation biology [e.g. @mace2008; @iucn2015]. These properties are key to everything from global assessments of extinction risk [e.g. @stuart2004; @schipper2008] to assessing the sustainability of hunting [e.g. @fa2002] and fishing exploitation levels [e.g. @worm2009; @juan-jorda2011]. However, conservation and ecological resource managers are often faced with reconciling multiple uncertain and potentially conflicting estimates of status and trajectory [e.g. @brodziak2010; @branch2011; @deroba2015]. Some models may suggest a population is at risk and declining in abundance while others may suggest it is not at risk and stable.

<!-- One solution is to... TODO KL: As part of setting the stage, right off the bat I'd delimit the set of conditions that make ensemble models an appropriate and useful choice, e.g.: When there are several models (how many to make it useful?) and there is a way to validate them, one solution is..."-->

@@

In fisheries science, the commonly used operational model for determining status and trajectory of exploited fish populations are stock assessments, i.e., population models coupled to an observation model, that incorporate all appropriate data (e.g. catches, size and age distributions, surveys, and tagging information) to quantify values such as the biomass of a stock that can produce maximum sustainable yield ($B/B_\mathrm{MSY}$) [@hilborn1992]. However, the broad range of data required to conduct these stock assessments are not available for the majority of fish populations, including those of conservation concern and of economic interest to fisheries [@fao2014]. Therefore, a number of models have been proposed to assess $B/B_\mathrm{MSY}$ based on the limited data available for the majority of fish stocks: (1) a time series of the total biomass of catch and (2) a basic understanding of population productivity [e.g. @vasconcellos2005; @martell2013]. Recently, @rosenberg2014 investigated the performance of four data-limited models through a large-scale simulation experiment. The four models frequently disagreed about population status and any one model had relatively poor predictive ability on average, but there was evidence suggesting that some models performed better than others depending on circumstances (e.g. Fig. \ref{motivate}).

Here, we estimate population status and trajectory of exploited fish populations using ensembles and superensembles (collectively 'ensemble methods') of these four data-limited models. We apply a variety of approaches to both simulated and real-world fish stocks and compare their predictive performance. *TODO Add a few sentences on the rationale for how the approaches were identified and the testing was designed.*

<!--We find that ensembles, and in particular superensembles, generally improved predictive performance of status and trajectory of fish populations over any single model---especially when considered across multiple dimensions of predictive ability. Furthermore, the output from superensembles can aid mechanistic understanding of how individual models of status and trajectory perform.-->

@@

We first developed and tested ensemble methods on a fully factorial simulated dataset [@rosenberg2014] with precisely known status. These published simulated data included the following scenarios: three fish life histories: small pelagic, demersal, and large pelagic; three levels of biomass depletion at the start of the dataset compared to carrying capacity: 100%, 70%, and 40% of carrying capacity; and four harvest dynamic settings: (1) a constant harvest rate, (2) a harvest rate coupled with biomass to mimic an open-access single-species fishery, (3) a scenario where harvest rate increases continuously, and (4) a 'roller-coaster' scenario where the harvest rate increases and then decreases. Process noise (i.e. variability in population dynamics) was introduced to the models through two levels of multiplicative recruitment variability: $N(0, 0.2)$ and $N(0, 0.6)$, and was either uncorrelated or had a first-order autoregressive correlation of $0.6$. The simulation also included a scenario with multiplicative observation error around catch at $N(0, 0.2)$ and one without observation error. 
<!--TODO: make last sentence clearer
4. First paragraph in Methods describing simulations line 113 in pdf - you may want to describe the recrutiment variability a bit more. I know its not an important detail, but because its in the first paragraph of methods, it might turn people off reading further. Most people wouldn't understand what 'multiplcative' (including me) is or may not know what 'process error/noise' refers to. 
5. First paragraph in Methods last sentence - needed a bit of rewording because as was, i got quite confused. hopefully this new sentence makes sense and doesnt change the meaning. 
-->

@@

We also tested our ensemble methods on the RAM Legacy Stock Assessment Database [@ricard2012]. Our analysis of the stock-assessment database was based on version XX downloaded on XX (TODO). After removing stocks for which the models described below did not converge (TODO: add number), this database included `r ram_stocks_n` stocks.
<!-- Dan: Do we address the implications of only running our method on stocks that converged later on? i.e. is there something about those omitted stocks that would make them less suited for this type of analysis?-->

@@

We fit four individual data-limited models to estimate $B/B_\mathrm{MSY}$. Three of the models are mechanistic and based generally on Schaefer (logistic) biomass dynamics [@schaefer1954] of the form

@@

\noindent
where $\hat{B}_{t+1}$ represents predicted biomass at time $t$ plus one year, $B_t$ represents biomass at time $t$, $r$ represents the intrinsic population growth rate, $B_0$ represents unfished biomass or carrying capacity $K$, and $C$ represents catch. The fourth model is an empirically derived model based on the RAM Legacy Stock Assessment Database. @rosenberg2014 provide a full summary of these four methods and code to fit all the models is available in an accompanying package `datalimited` (TODO: make public?) for the statistical software \textsf{R}. In summary:

* *CMSY* (catch-MSY) implements a stock-reduction analysis with Schaefer biomass dynamics [@martell2013]. It requires a prior distribution for $r$ and assigns a prior to the relative proportion of biomass at the end of the time series compared to unfished biomass (depletion) based on the percentage of maximum catch at the end of the time series. The version of the model used in @rosenberg2014 was modified from @martell2013 to generate biomass trends from all viable $r$-$K$ pairs and produce an estimate of $B/B_\mathrm{MSY}$ from the median trend.

* *COM-SIR* (catch-only-model with sampling-importance-resampling) is a coupled harvest-dynamics model [@vasconcellos2005]. Biomass is assumed to follow a Schaefer model and harvest dynamics are assumed to follow a logistic model. The model is fit with a sampling-importance-sampling algorithm [@rosenberg2014].

@@

* *mPRM* (modified panel regression model) is a modified version of the panel-regression model from @costello2012. Unlike the other models, mPRM is empirical and not mechanistic --- it uses the RAM Legacy Stock Assessment database to fit a regression model to a series of characteristics of the catch time series and stock with stock-assessed $B/B_\mathrm{MSY}$ as the response. The model used in this paper is modified from the original in that it condenses the life-history categories into three categories to match the simulated dataset, removes the maximum catch predictor since the absolute catch in the simulated dataset is arbitrary, and does not implement the bias correction needed in @costello2012 since we do not derive estimates of median status across multiple stocks.

@@

<!--TODO Kristin and Mark: It might be a good idea to add a figure that illustrates the spectrum of ensemble/superensembles that are applied here and what is included (e.g., non-linear relationships, etc.). This will help the reader understand the progression in complexity of the ensemble approaches I think.-->
Superensemble models allow us to incorporate additional covariates and potentially leverage interactions between these covariates and individual model predictions. Additional covariates could be, for example, life-history characteristics, information on exploitation patterns, or statistical properties of the data. For simplicity, and since many possible covariates, such as fishing effort dynamics, would not be known for many real-world data-limited scenarios, we added only one set of additional covariates: spectral properties of the catch time series. Spectral analysis decomposes a time series into the frequency domain and provides a means of describing the statistical properties of the catch series that is independent of time series length (except in affecting precision) and absolute magnitude of catch. We fit spectral models to the scaled catch time series (catch divided by maximum catch) with the `spec.ar` function in \textsf{R} and recorded representative short- and long-term spectral densities at frequencies of 0.20 and 0.05, which correspond to 5- and 20-year cycles. *TODO add new supp. fig.?*

<!-- TODO KL: I think it's worth explaining that we were interested in how the predictions changed wrt the shape of the real time-series, i.e. how their cyclic patterns of peaks and troughs matched, and that's why other summary statistics, dependent on abundance or length but insensitive to the shape of the trajectory, weren't chosen.-->

@@

The individual models we seek to combine with superensembles provide time series of stock status ($B/B_\mathrm{MSY}$). Therefore, we can use superensembles to estimate multiple properties of these time series. Here, we focus on two properties: the mean and slope of $B/B_\mathrm{MSY}$ in the last 5 years. Together, these quantities address the recent state and trajectory of stock status, which are both of management and conservation interest [e.g. @mace2008]. To avoid undue influence of the end points of the time series on the calculated slope, we measured the slope as the Theil-Sen estimator of median slope [@theil1950].

@@

A critical component to any predictive modelling exercise is to evaluate the performance of a model on new data [@hastie2009]. We used repeated three-fold cross validation to test predictive performance: we randomly divided the dataset into three sets, built superensemble models on two-thirds of the data, and evaluated predictive performance on the remaining third. We repeated this across each of the three splits and then repeated the whole procedure 50 times to account for bias that may result from any one set of validation splits. In the simulated dataset, there were 10 replicates of each unique combination of simulation parameters that differed only in stochastic variability. Since the dynamics of these populations were often similar, we grouped these stocks in the cross-validation process into either the training or testing split. When applied to the RAM Legacy Stock Assessment database, we used superensembles trained on the entire simulation dataset. However, since mPRM is built on the same stock assessment database, we refit mPRM on each training split.

Predictive performance can be evaluated with metrics that represent a variety of modelling goals. For continuous response variables such as the mean and slope of population status, performance metrics often measure some form of bias, precision, accuracy (a combination of bias and precision), or the ability to correctly rank or correlate across populations [e.g. @walther2005]. Here, we measure proportional error, defined as $(\hat{b} - b)/b$, where $\hat{b}$ and $b$ represent estimated and 'true' (or stock-assessed) mean or slope of $B/B_\mathrm{MSY}$. We calculated median proportional error to measure bias, median absolute proportional error to measure accuracy, and Spearman's rank-order correlation between predicted and 'true' values to measure the ability to correctly rank populations. We measured the ability for each model to correctly place stocks into categories of above or below $B/B_\mathrm{MSY} = 1$ at any decision threshold by plotting receiver operating characteristic curves. Code to reproduce our analysis is available at <https://github.com/datalimited/ensembles> (TODO: make public). When cross-validating against the RAM Legacy Stock Assessment database, we treat the estimates from these data-rich stock assessments as known without error. Thus, any error in the stock assessment estimates of the mean or slope of $B/B_\mathrm{MSY}$ also contributes to our estimates of prediction error for each of the four data-limited models and the ensembles.

*TODO Olaf: B/Bmsy = 0.5 may be a better threshold as this corresponds to the overfished definition in the US and a few other places.  As Ray always points out, if Bmsy is the target, then you’d expect to be below it about half of the time in a well managed fishery.*

<!-- TODO PE .... mean or slope of b/bmsy: Dan: logged or normal space? For the status regression where you are estimating log(B/Bmsy), wouldn't you need to correct for the retransformation bias in the estimating median proportional error later down? I could be missing something though in whether that applied once you've calculated proportional error-->

<!-- TODO Dan: Should we discuss the issue with MAPE that we ran into before, in that something can have bimodal predictions that are really positively or really negatively biased, and the net result its a MAPE of 0?-->

@@

Applied to the simulated dataset of known status, the individual models had variable success at recovering the mean (status) and slope (trajectory) of $B/B_\mathrm{MSY}$ in the last five years. All models exhibited a high degree of scatter around the one-to-one line of perfect status prediction (Fig. \ref{hexagon}). In contrast to the known unimodal distribution of status, CMSY exhibited bimodal predictions (Fig. \ref{hexagon}a), but had the best rank-order correlation and accuracy scores (Fig. \ref{performance}a). COM-SIR and SSCOM both correctly identified a number of stocks with low status, but frequently predicted a high status when status was in fact low (Fig. \ref{hexagon}b, c). mPRM had relatively poor ability to predict status for the simulated dataset (Fig. \ref{hexagon}d). There was generally little correlation between true and predicted trajectory for any of the individual models (rank-order correlation = `r range_slope_corr_non_sscom`) with the exception of SSCOM (correlation = `r slope_corr_sscom`; Figs \ref{scatter-sim-slope}a--d).

@@

<!-- TODO KMK: "by exploiting interactions between the spectral density predictors and the individual model predictions" I'm not exactly sure I follow this and I suspect that the subtleties would be lost on others. Can you explain this a little more? -->

@@

<!-- TODO Dan: worth noting in the discussion that this is still pretty low: These methods clearly improve the predictive ability of the catch-based DLAS, but we should point out that it hardly moves it to a perfect estimator; there's still a lot of error in any individual estimate-->

@@

<!-- Dan:
A few general comments. Great job, this is well written and clear, especially given the esoteric nature of the content. In the discussion, can we address actual limitations of ensemble methods, if any? We say when it works, when might it not? Are there times when it would actually be a bad idea to use ensembles, or would it always be as good or better than indivduals or simple means?

Second, it might be worth mentioning that the ensembles here are helping make sense of conflicting results from the same data. Do these methods work if you were trying to compare methods each individually estimated using different data? This same method could presumably work for any metric or data, but it's worth noting that we've only tested catch-based methods here.

Lastly, it's worth noting that this approach, as laid out here is for helping improve potentially conflicting estimates of the same metric, B/Bmsy. In many cases though a fishery is going to be basing its decisions on streams of multiple indicators, e.g trends in SPR and B/Bmsy. While the ensemble methods here can help refine estimates of individual metrics, it's still worth looking at what alternative data streams are telling you if possible. E.g. if a purely catch based B/Bmsy estimate is high and has a positive slope, but mean lengths and SPR are decreasing, managers will still need to think about how to deal with these conflicting signals. So, this ensemble method, as I interpret it, is helpful for resolving conflicts within metric, but doesn't necesarily resolve conflicts across metrics (if I'm understanding these methods right)
-->

@@

<!-- TODO KMK: Does it make a difference that 3 of our models are based on Schaefer? Definitely they have different structural forms, but not sure if this warrants any discussion?-->

Certain conditions will make some ensemble models more effective than others. First, ensembles will be most effective when they are comprised of diverse individual models that choose different structural model forms, explore contrasting but plausible ranges of parameter values, or make uncorrelated errors [@ali1996; @dietterich2000; @tebaldi2007]. Such individual models would be expected to perform well in different conditions and an ensemble model can exploit the best predictive performance of each. Second, ensemble models will be most effective when they are not overfit to the training dataset. Cross-validation testing [@caruana2004; @hastie2009] and methods that are robust to overfitting such as random forests [@breiman2001], may help avoid overfitting ensemble models. We note that our simplest ensemble model, an average of individual model predictions, performed approximately as well as complex machine learning models when we trained our superensembles on the simulation dataset and tested them on a separate 'real' dataset (i.e. the RAM Legacy Stock Assessment database, Fig. \ref{performance}b). Third, ensemble models will be most effective when they are trained on data that are representative of the dataset of interest [@knutti2009; @weigel2010]. Cross-validation within a training dataset will provide an optimistically biased impression of predictive performance if the training dataset fundamentally differs from the dataset of interest [@hastie2009].

Multi-model inference in the form of coefficient averaging weighted by information theoretics such as the Akaike Information Criterion (AIC) is a common analytical approach in ecology [@burnham2002; @johnson2004; @grueber2011]. The ensemble methods described in this paper share similarities with coefficient averaging but differ in other important ways. Ensemble methods and coefficient averaging share the long-held notion that multiple working hypotheses can contribute useful information for inference [@chamberlin1890]. A fundamental difference is that coefficient averaging focuses on averaging *coefficients* whereas ensembles instead average *predictions*. Thus, while the outcome will be the same in some cases, ensembles provide a more general purpose tool: they do not require information theoretics and they can combine different types of models (e.g. parametric and non-parametric models or frequentist and Bayesian predictions). Furthermore, superensembles extend these benefits by allowing model predictions to be combined via non-linear functions that are tuned to known data.
